---
title: "Pre-processing"
author: "Rochelle Terman"
date: "June 24, 2015"
output: html_document
---

### Required Packages

First load our required packages:
```{r}
setwd("~/Dropbox/berkeley/PS239T/text-analysis")
rm(list=ls())
library(tm) # Framework for text mining
library(RTextTools) # a machine learning package for text classification written in R
library(qdap) # Quantiative discourse analysis of transcripts
library(qdapDictionaries)
library(dplyr) # Data preparation and pipes $>$
library(ggplot2) # for plotting word frequencies
library(SnowballC) # for stemming
```

## 1. Prepare a Corpus

A corpus is a collection of texts, usually stored electronically, and from which we perform our analysis. A corpus might be a collection of news articles from Reuters or the published works of Shakespeare. Within each corpus we will have separate articles, stories, volumes, each treated as a separate entity or record.

Documents come in a variety of formats, but plain text is best.

### 1.1 Corpus Sources and Readers

The `tm` package supports a variety of sources and formats. 

```{r}
getSources()
getReaders()
```

We can read a corpus from a directory that contains text files, each document a different file
```{r}
documents <- Corpus(DirSource("Data/Sessions")) 
documents
```

Or we can read from a csv of documents, with each row being a document, and columns for metadata (information about each document)

```{r}
documents <-read.csv("Data/shelbysessions.csv", header=TRUE) #read in CSV file
docs <- Corpus(VectorSource(documents$text))
docs
```

We can inspect the documents using inspect()

```{r}
inspect(docs[16])
```
### 1.2 Preprocessing

Many text analysis applications follow a similar 'recipe' for preprecessing, involving:

1. Tokenizing the text to unigrams (or bigrams, or trigrams)
2. Converting all characters to lowercase
3. Removing punctuation
4. Removing numbers
5. Removing Stop Words, inclugind custom stop words
6. "Stemming" words, or lemmitization. There are several stemming alogrithms. Porter is the most popular.
7. Removing Sparse Terms
8. Creating a Document-Term Matrix

See what transformations are available TM package

```{r}
getTransformations()
```

The function `tm.map()` is used to apply one of these transformations across all documents. 

We can also use regular R functions for custom transformations using the `content_transformer()` function.

For example, we might want to replace “/”, used sometimes to separate alternative words, with a space. This will avoid the two words being run into one string of characters through the transformations. We might also replace “@” and “|” with a space, for the same reason
```{r}
# note: won't work on past versions of R
toSpace <- content_transformer(function(x, pattern) gsub(pattern, " ", x))
docs <- tm_map(docs, toSpace, "/")
docs <- tm_map(docs, toSpace, "@")
docs <- tm_map(docs, toSpace,"\\|")
docs[[100]]
```

We can basic transformations using the build in `tm_map` tools.

```{r}
docs <- tm_map(docs, content_transformer(tolower)) # convert all text to lower case
inspect(docs[16])
docs <- tm_map(docs, removePunctuation) # remove Puncturation
inspect(docs[16])
docs <- tm_map(docs, removeNumbers) # remove Numbers
inspect(docs[16])
docs <- tm_map(docs, removeWords, stopwords("english")) # remove common words
stopwords("english") # check out what was removed
inspect(docs[16])
docs <- tm_map(docs, removeWords, c("shelby","sessions")) # remove own stop words
inspect(docs[16])
docs <- tm_map(docs, stripWhitespace) # strip white space
inspect(docs[16])
docs <- tm_map(docs, stemDocument) # stem the document
inspect(docs[16])
```

### 1.3 Creating a DTM

A document term matrix is simply a matrix with documents as the rows and terms as the columns and a count of the frequency of words as the cells of the matrix. We use `DocumentTermMatrix()` to create the matrix:
```{r}
dtm <- DocumentTermMatrix(docs)
dtm
dim(dtm) # how many documents? how many terms?
inspect(dtm[1:5,1000:1005]) # take a quick look
```

Note: we can use RTextTools to go directly from text vector to DTM, which is probably easier.

```{r}
# make DTM
dtm <- create_matrix(docs[["text"]], language="english", removeNumbers=TRUE,
                       stemWords=TRUE, removeSparseTerms=.9, toLower = TRUE, 
                       removePunctuation = TRUE)
```

### 1.4 Exploring the DTM

We can obtain the term frequencies as a vector by converting the document term matrix into a matrix and summing the column counts:

```{r}
freq <- colSums(as.matrix(dtm))
length(freq) # how many terms?
```

By ordering the frequencies we can list the most frequent terms and the least frequent terms:

```{r}
# order
ord <- order(freq)

# Least frequent terms
freq[head(ord)]

# most frequent
freq[tail(ord)]

# frequency of frenquencies
head(table(freq),15)
tail(table(freq),15)

# plot
plot(table(freq))
```
Exploring word frequences

```{r}
# Have a look at common words
findFreqTerms(dtm, lowfreq=1000) # words that appear at least 1000 times

# Which words correlate with "congress"?
findAssocs(dtm, "fire", 0.7)

# plot
freq <- sort(colSums(as.matrix(dtm)),decreasing=TRUE)
head(freq)

wf <- data.frame(word=names(freq), freq=freq)
head(wf)

library(ggplot2)
subset(wf, freq>1000) %>%
  ggplot (aes(word, freq)) +
  geom_bar (stat ="identity") +
  theme(axis.text.x=element_text(angle=45,hjust=1))

# wordcoulds!
library(wordcloud)
set.seed(123)
wordcloud(names(freq), freq, max.words=100, colors=brewer.pal(6,"Dark2"))
```
We can remove sparse terms and thus inrease efficency

```{r}
dtm.s <- removeSparseTerms(dtm,.99)
dtm # 6506 terms
dtm.s # 1849 terms
```

### 1.5 Exporting the DTM

We can convert a DTM to a matrix or data.frame in order to write to a csv, add meta data, etc.

```{r}
# coerce into dataframe
dtm <- as.data.frame(as.matrix(dtm))
names(docs)

# add column for author
dtm$document_author <- documents$author

# check to see if they're the same number of documents per author
summary(dtm$document_author)
summary(documents$author)
```